---
output:
  word_document: default
  html_document: default
---
# ClassTreeAssign1 
## Mod 4 Assin 1
### Evan Hildreth 

```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
```
```{r}
heart_disease <- read_csv("heart_disease-1.csv")
```

```{r}
heart_disease <- heart_disease %>% mutate(Sex = as_factor(Sex)) %>% 
  mutate(ChestPainType = as_factor(ChestPainType)) %>% 
  mutate(RestingECG = as_factor(RestingECG)) %>% 
  mutate(ExerciseAngina = as_factor(ExerciseAngina)) %>% 
  mutate(HeartDisease = as_factor(HeartDisease)) %>% 
  mutate(HeartDisease = fct_recode(HeartDisease, "No" = "0", "Yes" = "1"))

```

### Task 1 

```{r}
set.seed(12345)
tearing_my_heart_up <- initial_split(heart_disease, prop = 0.7, strata = HeartDisease)
heart_disease_training <- training(tearing_my_heart_up)
heart_disease_testing <- testing(tearing_my_heart_up)

```

### Task 2

```{r}
heart_disease_recipe1 <- recipe(HeartDisease ~ ., heart_disease)

Tree_model <- decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

heart_disease_wflow1 <- 
  workflow() %>% 
  add_model(Tree_model) %>% 
  add_recipe(heart_disease_recipe1)

heart_disease_fit <- fit(heart_disease_wflow1, heart_disease_training)

heart_disease_pluck <- heart_disease_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(heart_disease_pluck)
```

### Task 3

```{r}
heart_disease_fit$fit$fit$fit$cptable

```

The CP value that corresponds with the smallest xerror value (0.3588850) is 0.01742160. 

### Task 4 
```{r}
set.seed(123)
folds <- vfold_cv(heart_disease_training, v = 5)

heart_disease_recipe2 <- recipe(HeartDisease ~., heart_disease) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model2 <- decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")


heart_disease_grid <- grid_regular(cost_complexity(), levels = 25)
                          
heart_disease_wflow2 <- 
  workflow() %>% 
  add_model(tree_model2) %>% 
  add_recipe(heart_disease_recipe2)

heart_disease_res <- 
  heart_disease_wflow2 %>% 
  tune_grid(
    resamples = folds,
    grid = heart_disease_grid
    )

heart_disease_res
```

```{r}
heart_disease_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 

```
### Task 5
Looking at the above graphs it appears that CP value ~ 0.04 looks to provide the optimal accuracy value. 

### Task 6 

```{r}
best_tree_HDR <- heart_disease_res %>%
  select_best("accuracy")

best_tree_HDR

final_wf <- 
  heart_disease_wflow2 %>% 
  finalize_workflow(best_tree_HDR)

final_fit_HDR = fit(final_wf, heart_disease_training)

tree_HDR = final_fit_HDR %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(tree_HDR, tweak = 1.5) 

```

### Task 7 

```{r}
HDR_predict <- predict(final_fit_HDR, heart_disease_training, type = "class")
head(HDR_predict)
confusionMatrix(HDR_predict$.pred_class, heart_disease_training$HeartDisease, positive = "Yes")
```
Running a confusion matrix on the mode (prior to it being plucked) we find that the model has an accuracy of 0.8396. 

### Task 8 
```{r}
Blood <- read_csv("Blood.csv")
```

```{r}
Blood <- Blood %>% mutate(DonatedMarch = as_factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1"))

```

### Task 9 

```{r}
set.seed(1234)
Blood_spilt <- initial_split(Blood, prop = 0.7, strata = DonatedMarch)
Blood_training <- training(Blood_spilt)
Blood_testing <- testing(Blood_spilt)

```

```{r}
set.seed(1234)
blood_folds <- vfold_cv(Blood_training, v = 5)

blood_recipe <- recipe(DonatedMarch ~., Blood) %>%
  step_dummy(all_nominal(),-all_outcomes())

blood_model <- decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")


blood_grid <- grid_regular(cost_complexity(), levels = 25)
                          
blood_wflow <- 
  workflow() %>% 
  add_model(blood_model) %>% 
  add_recipe(blood_recipe)

Blood_res <- 
  blood_wflow %>% 
  tune_grid(
    resamples = blood_folds,
    grid = blood_grid
    )

Blood_res

```

```{r}
Blood_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```
Looking at the chart above, it seems the cp value of roughly ~0.020 possesses the highest accuracy rating.  

```{r}
best_bloody_tree <- Blood_res %>%
  select_best("accuracy")

best_bloody_tree

Bloody_final_wf <- 
  blood_wflow %>% 
  finalize_workflow(best_bloody_tree)

bloody_fit = fit(Bloody_final_wf, Blood_training)

Bloody_tree_fit = bloody_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(Bloody_tree_fit, tweak = 1.5) 
```
### Task 11 

```{r}
Blood_predict <- predict(bloody_fit, Blood_training, type = "class")
head(Blood_predict)
confusionMatrix(Blood_predict$.pred_class, Blood_training$DonatedMarch, positive = "Yes")
```
With the blood dataset and the blood model (essentially the same as the tree_model2 just renamed) we get an accuracy of 0.8069 which is slightly lower than the accuracy received on the heart_disease dataset (0.8396 accuracy). 

In addition, while the P value for both datasets with the model is significant, it is much better on heart disease dataset than the blood one. 